---
title: "Social Network Analysis of the Malaria Co-authorship network from Benin"
output: html_notebook
author: "Roseric Azondekon"
---
$\\$
$\\$

This project is part of my PhD dissertation project.

This section follows from the first section dedicated to the Stochastic Block Modeling of our co-authorship network (available [HERE](http://#)).

In this section, we apply Dynamic Stochastic Block Modeling to our data.

Loading required packages...
```{r}
library(mixer)
library(igraph)
library(doParallel)
library(parallel)
```

Setting Working directory and setting seeds...
```{r}
setwd('~/R/R_scripts')
set.seed(123)
```

Preparing Temporal network data...
```{r}
# Loading a simplified version of the network
mnet <- read_graph('./graphs/CAnet.graphml', format = 'graphml')

# mnet2.96.01 <- read_graph('./graphs/CAnet_96_01s.graphml', format = 'graphml')
# mnet2.02.06 <- read_graph('./graphs/CAnet_02_06s.graphml', format = 'graphml')
# mnet2.07.11 <- read_graph('./graphs/CAnet_07_11s.graphml', format = 'graphml')
# mnet2.12.16 <- read_graph('./graphs/CAnet_12_16s.graphml', format = 'graphml')
```


```{r}
# Aggregating network by time periods
# mnet2.96.01 <- subgraph.edges(mnet,which(E(mnet)$year<=2001), delete.vertices = F)
mnet2.96.01 <- subgraph.edges(mnet,which(E(mnet)$year<=2001), delete.vertices = F)


mnet2.02.06 <- subgraph.edges(mnet,which(E(mnet)$year<=2006 & E(mnet)$year>=2002), delete.vertices = F)


mnet2.07.11 <- subgraph.edges(mnet,which(E(mnet)$year<=2011 & E(mnet)$year>=2007), delete.vertices = F)


mnet2.12.16 <- subgraph.edges(mnet,which(E(mnet)$year>=2012), delete.vertices = F)


V(mnet2.96.01)$name <- as.character(1:vcount(mnet2.96.01))
V(mnet2.02.06)$name <- as.character(1:vcount(mnet2.96.01))
V(mnet2.07.11)$name <- as.character(1:vcount(mnet2.96.01))
V(mnet2.12.16)$name <- as.character(1:vcount(mnet2.96.01))

list.mnet <- list(mnet2.96.01, mnet2.02.06, mnet2.07.11, mnet2.12.16)
```

```{r}
E(mnet)$weight <- 1
mnets <- simplify(mnet, edge.attr.comb = list(weight="sum"))
```


Converting to array of adjacency matrices
```{r,warning=FALSE}
# converting to R array
all.vertices.name <- unique(sort(sapply(list.mnet, function(g) V(g)$name)))
T <- length(list.mnet)
N <- length(all.vertices.name)
Y <- array(dim=c(T,N,N))
for (t in 1:T){
   g.tmp <- graph.empty(n=N, directed=F)
   V(g.tmp)$name <- all.vertices.name
   Y[t,,] <- as.matrix(get.adjacency(str(g.tmp %u% list.mnet[[t]]), sparse = F))
}
```


Fitting Dynamic SBM...
```{r}
# dynSBM
t0 <- Sys.time()
list.dynsbm <- select.dynsbm(Y, Qmin=2, Qmax=5, edge.type="binary", nstart = 1)
Sys.time() - t0
dynsbm <- list.dynsbm[[2]]
# Connectivity Plot
connectivity.plot(dynsbm, Y)
```

```{r}
p<-alluvial.plot(dynsbm)
```

