---
title: "Social Network Analysis of the Malaria Co-authorship network from Benin"
output: html_notebook
author: "Roseric Azondekon"
---
$\\$
$\\$

This project is part of my PhD dissertation project.

This section follows from the first section dedicated to the Mathematical Modeling of our co-authorship network (available [HERE](http://#)).

The purpose of this section is the use of Exponential Random Graph Modeling (ERGM) as a  Statistical modeling to better understand and explain our malaria co-authorship network.

Loading the necessary packages...
```{r}
setwd('~/R/R_scripts')
library(igraph)
library(ergm)
library(parallel)
library(doParallel)
library(lubridate)
# library(Rmpi)
library(snow)
library(methods)
```

Loading the variables from the last section...
```{r}
load('./Rdata/mnet.rda')
load('./Rdata/mnet2.rda')
load('./Rdata/mnet2.gc.rda')
load('./Rdata/auth_data.rda')
load('./Rdata/edges.rda')
load('./Rdata/mnet2.gc.rda')

# Loading a simplified version of the network
mnet3 <- read_graph('./graphs/CAnet_graph.graphml', format = 'graphml')
```

Since this processing is going to take a lot of time, we decide to parallelize all the computations:
```{r}
# # create a cluster with fixed number of processes
# np = detectCores()-2 # Detectin CPUs and leaving2 out
# cluster <- makePSOCKcluster(np)

cores=detectCores() # get number of cores
cl <- makeCluster(cores[1]-2) #I decide to run leave 2 cores out
registerDoParallel(cl)
```

```{r}
A <- get.adjacency(mnet3)
v.attrs <- get.data.frame(mnet3, what="vertices")

v.attrs$name <- V(mnet2)$name
v.attrs$timesCited <- V(mnet2)$timesCited
v.attrs$numPub <- V(mnet2)$numPub
v.attrs$community <- V(mnet2)$community
v.attrs$degree <- V(mnet2)$degree
v.attrs$affiliation <- V(mnet2)$place
v.attrs$city <- V(mnet2)$affil
v.attrs$country <- V(mnet2)$country
```

```{r}
mnet3.simple <- network::as.network(as.matrix(A), directed=FALSE)
network::set.vertex.attribute(mnet3.simple, "timesCited", V(mnet2)$timesCited)
network::set.vertex.attribute(mnet3.simple, "numPub", V(mnet2)$numPub)
network::set.vertex.attribute(mnet3.simple, "community", V(mnet2)$community)
network::set.vertex.attribute(mnet3.simple, "degree", V(mnet2)$degree)
```



```{r}
my.ergm.bern <- formula(mnet3.simple ~ edges)
my.ergm.bern
mnet3.simple ~ edges
summary.statistics(my.ergm.bern)
```

```{r}
t0 <- Sys.time()
my.ergm.bern.fit <- ergm(my.ergm.bern, control=control.ergm(parallel=cores, parallel.type="PSOCK", MCMC.interval=1024*10,  MCMC.burnin=1024*16, MCMC.samplesize=1024*10))
Sys.time() - t0
```

```{r}
anova.ergm(my.ergm.bern.fit)
```

```{r}
summary(my.ergm.bern.fit)
```

```{r}
t0 <- Sys.time()
diag <- gof(my.ergm.bern.fit, control=control.gof.ergm(parallel=cores, parallel.type="PSOCK", MCMC.interval=1024*10))
Sys.time() - t0
```

```{r}
# save(diag, file = './Rdata/diag.rda')
```


```{r,fig.width=20,fig.height=15}
par(mfrow=c(1, 4))
plot(diag) # Not a quite good fit.
```



```{r}
my.ergm <- formula(mnet3.simple ~ edges + kstar(2) + kstar(3) + triangle)
summary.statistics(my.ergm)
```

```{r}
t0 = proc.time()
my.ergm.fit <- ergm(my.ergm, control=control.ergm(parallel=cores, parallel.type="PSOCK", MCMC.interval=1024*10,  MCMC.burnin=MCMC.interval*16, MCMC.samplesize=1024*10))
proc.time() - t0
```

```{r}
anova.ergm(my.ergm.fit)
summary(my.ergm.fit)
```

```{r}
mcmc.diagnostics(my.ergm.fit) # to check for diagnostics
```


```{r}
my.ergm <- formula(mnet3.simple ~ edges + gwesp(1, fixed=TRUE))
summary.statistics(my.ergm)
```

```{r}
mnet3.ergm <- formula(mnet3.simple ~ edges 
                       + gwesp(log(3), fixed=TRUE)
                       + nodemain("timesCited") 
                       + nodemain("numPub")
                       + nodemain("degree") 
                       + match("community"))
```

```{r}
t0 = proc.time()
mnet3.ergm.fit <- ergm(mnet3.ergm, control=control.ergm(parallel=cores, parallel.type="PSOCK"))
proc.time() - t0
```

```{r}
mcmc.diagnostics(mnet3.ergm.fit) # to check for diagnostics
```

```{r}
anova.ergm(mnet3.ergm.fit)
```

```{r}
summary.ergm(mnet3.ergm.fit)
```

```{r}
gof.mnet3.ergm <- gof(mnet3.ergm.fit,model=F)
```

```{r}
plot(gof.mnet3.ergm)
par(mfrow=c(1, 4))
plot(gof.mnet3.ergm)
```






```{r}
stopCluster(cl)
```

